Parallel multi-objective optimization using selective operator application islands

### 2017-07-03 hpmoon-coor.tex
Addressing more changes by jesus, julio and @jj

\begin{abstract}
  % First, state the problem!!!! - JJ Pablo: ok!
Real-world optimization problems usually require to deal with several objectives.
Multi-objective optimization methods have gained interest in the last
years because of the performance they can achieve solving this kind of problems. 
Many of them usually require a high-dimensional
decision space, and multi-objective Evolutionary Algorithms have been
successfully used in the past. Due to the large computational cost
these algorithms require, different parallelization and distribution
methods have been proposed, with different parallel and distributed
computer architectures to cope with them.

One of the most extended
methods are the distributed co-evolutionary algorithms, where each
processor deals only with a part of the decision space, that is, the
individuals belonging to the same sub-population (island) explore the
same subset of decision variables. However, these methods only works with decomposable problems.

A similar idea is presented here to deal with un-decomposable problems, the selective operator application:
instead of sharing parts of individuals, each processor applies the variation 
operators to a specific subset of the whole individual. We have previously proved 
that extending the length of these subsets, and therefore, sharing some search 
space between islands, implies an improvement in the results.

In this paper, a new method to
automatically adapt the size of the overlapping, or shared sections,
of the individuals has been compared to other collaborative
evolutionary techniques. Different number of islands, chromosome sizes
and problems have been used. The analysis of the obtained experimental
results, by using different metrics, shows that selective operator application
approaches can provide statistically significant improvements with
respect to the base algorithm, being the method to automatically adapt
the overlapping size the one to obtain the best performance during the
same execution time. Results also show that the relation of the number
of islands (subpopulations) to the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables. 
\end{abstract}

@@

%2 COMPUTATION -> PARALLEL
These kind of problems require a high demand of computational time, leading to the apparition of easily parallelizable methods to solve them \citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress}.
%3 ONE WAY IS TO USE MOEAS BECAUSE
One example of easily parallelizable algorithms to solve MOPs are the Multi-Objective Evolutionary Algorithms (MOEAs), a sub-set of Evolutionary algorithms (EAs). EAs \citep{DBLP:series/ncs/EibenS15} are bio-inspired meta-heuristics that can be effectively used to find nearly optimal solutions for optimization problems. Usually an EA starts by generating a set of random solutions, called \emph{population}, following a user-defined description. Then, it evaluates each candidate solution, called \emph{individual}, assigning it a \emph{fitness} value, that describes how good the individual is, with regard to the target problem. New solutions are then generated by the application of \emph{operators} that either mutate a single solution or recombine different existing solutions. After each iteration, called \emph{generation}, the least fit individuals are removed, and the process continues until a user-defined stop condition is met. One of the most well-known and referenced MOEAs \citep{Dorronsoro13superlinear} is the {\em
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, and divided into different ranks (based on the
dominance) to be selected for the next generations.

@@

Different approaches have been used to parallelize EAs, as each individual can be considered as an
independent unit \citep{Alba13parallel}. Classic methods, such as the
global parallel EAs (Master-Slave), or the spatially structured
algorithms (Island model or Cellular EAs) have been applied
successfully in the past \citep{Folino03cellular,Alba02Parallelism}.
However, in the case of MOEAs, these
approaches \citep{Luna15Survey} need to deal with the whole solution set, the PF.
 This imply to use different distribution
and sharing mechanisms, as there exist a tension between the speedup
achievable from parallelization and the need to globally recombine the
results to accurately identify the PF
\citep{Branke04Parallelizingcone}.

MOPs usually require a high number of variables, meaning that the MOEAs need to deal with large individuals and spend significant extra time for crossover, mutation and migration. Different authors have proposed methods to divide the decision space (the chromosome) to improve the performance and solution quality. In this aspect, the co-evolution model is a dimension-distributed model where a high-dimensional problem is divided into lower dimensional problems \citep{Gong15models,Tonda12cooperative}, and evolved separately. One example of application of this technique was described in \citep{Kimovski15Parallel}. The method presented in that works involves different workers that evolve sub-populations created and recombined by a master process, which performs different recombination alternatives from the parts returned by worker processes. A high dimensional problem was used to compare these
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

The hypothesis of this paper is that using overlapped sections of the
chromosome for the SLA, and obtaining the number of sections automatically from the number of available islands, in a
multi-objective island-based algorithm, in a real computational environment, can improve the
quality of the solutions in the same computing time with respect to
the baseline, disjoint and partly overlapped methods. %That was the
                                %hypothesis of the previous
                                %paper. This paper has a new
                                %hypothesis, right? - JJ FERGU: no, this is the hypothesis of this paper (comparing adaptive with partly overlapped), but clarifying it
In order to prove this claim, our new adaptive method has been
compared against different overlapping islands schemes: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
SLA version with disjoint sections and a SLA
fixed-size overlapped version. %What was done in the previous paper?
                               %Are we re-using the results in that
                               %paper? - JJ FERGU: explained before that we are showing new results.
Different benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.  % Shouldn't this start with "We are trying to
                        % prove..." - JJ FERGU: Already writed in previous paragraph

@@

% Is this really cooperative co-evolution? - JJ FERGU: You mean the Dorronsoro one? The title says it
Addressing high-dimensional problems with cooperative co-evolution has
also been studied in several works with approaches closer to the one
presented here. The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. This approach has also been used by Kimovski et al. \citep{Kimovski15Parallel}, but using a
master-slave method that splits the population into several processors. As in previous work, each node
runs a parallel MOEA that only affects some portion of 
the individuals and the master
process receives all the sub-populations to be combined every certain number of generations. Up to 8 processors
were used in the experimental setup, and several
combination alternatives were compared. The main difference of our work with respect to previous works is that our approach does not 
broadcast all solutions to all islands for recombination, but only one solution to a random island, needing less
communication time. Moreover, Dorronsoro or Kimovsky approaches limited the maximum number of islands to 8, while in this paper we have
used up to 128 islands.

@@

In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two different dMOEAs. The first
one divided the chromosome in $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivated us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and this time, the experiments have been run on a real
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, and also validating the
previous methods.

@@

%This probably should deserve a "Algorithm" environment too - JJ FERGU: You mean the LaTeX one? I've used it in the pseudocode figures
This basic distributed algorithms spread the population among $P$
islands; after a fixed number of generations,
one individual in a given island is migrated to another random island,
thus avoiding to synchronize the global Pareto Front (PF) every
certain number of generations as other methods described in Section 2
do. At the end of the run, the Pareto Fronts of all islands are
aggregated into a new one, and the quality measures are evaluated. 
Different  SLA alternatives can be devised to evolve the
subpopulations according to the decision space to be explored by each
islands. Here we will consider three methods in which each island only
performs crossover and mutation in specific sections of the
individuals.

@@

The criterion used for terminating an experiment has been the running
time: 25 seconds for dimension 512 and 100s (four times more) for
2048. We have used time instead of the number of evaluations firstly
because our hypothesis argues that the time saved in crossover and
mutation can be spent on improving the sub-populations and more
operations and migrations can be achieved. Also, we are using
different number of islands (with different sub-population sizes) and that
could lead to different execution times, so it would be difficult to
compare different times and quality of solutions at the same time.

@@

Different metrics, explained in the previous section, have been used
to calculate the quality of the obtained PFs in each configuration. As
some of the metrics  (such as HV) require a reference point to be
calculated, the point (1,9) has been chosen as reference, as none of
the generated PFs in all runs are dominated by it. Metrics are then
normalized with respect to that point. A Kruskal-Wallis significance
test has been performed to the metrics of all runs of the
configurations, as the Kolmogorov-Smirnov test detected non-normal
distributions. The average results for each configuration are shown in
Table \ref{tab:results512} (for 512 dimensions) and
\ref{tab:results2048} (for 2048). As previously explained, when $P=8$
the results of A are equivalent to the obtained with D (because
$c=0$), and when $P=16$ they are equivalent to O ($c=1$).

@@

%DESCRIPTION OF THE RESULTS
Results show that there is a difference in performance when using 512 or 2048 dimensions. Although in some cases the overlapped methods obtain quality metrics values better or significantly equal than the obtained by the baseline, it is clear that 512 dimensions is not a value high enough to apply these methods. This contradicts the results obtained in our previous work in a shared-memory mono-processor version \citep{Garcia16hpmoon}, where the overlapping method obtained better results than the baseline in both dimension lengths. This can be explained because the added migration latency between nodes of the cluster requires more time than the time saved during the crossover and migration. This was not a problem in the mono-processor version, where the individuals shared the same memory space.

@@

%EXPLANATION OF THE RESULTS
This can be explained comparing the number of non-dominated solutions
in each island and the average number of generations (Table
\ref{tab:sols512} and Table \ref{tab:sols2048}). With both chromosome
lengths, increasing the number of islands implies more generations
with all the methods (logically, as there are less individuals in each
island). But also, the SLA methods
approach to the number of generations with the baseline when
increasing the number of islands. However, with $L=2048$ the migration
time is enough big to almost not improve the number of generations
with respect to the baseline, even in the lower number of islands, as
it happens with $L=512$. Therefore, more generations do not
necessarily improve the solution of the global PF, but focusing on
different elements of the chromosome. As indicated previously, every
island is unaware of the other islands PFs, and they are trying to
optimize their solutions independently. With respect to the average
number of solutions in each island, there is a clear difference in
number with the baseline with $L=512$, where this value is in the most
of the cases, less than a half. The number of non-dominated solutions
also improves a better Spread indicator, where the baseline almost
always obtains better values. %TENGO QUE COMENTAR MAS LOS RESULTADOS Y
                              %ENTENDERLO TODO Y JUSTIFICARLO
                              %FINAMENTE
%Almost always no es muy preciso - JJ

@@

Problems that require high-performance % problems are not high-performance. They
                          % _require_ high performance Fergu: done
 and deal with a large number of decision variables can leverage the
 division of the decision space that parallel and distributed
 algorithms imply. This can be done in dMOEAs by selective operator application (SLA), that is, dividing the
 chromosome into different parts, each one modified by a different
 island. This paper compares a baseline distributed NSGA-II with three
 different strategies to separate the chromosome (disjoint or
 overlapping parts), using a high number of islands. Results show that
 these methods can achieve better quality metrics than the baseline in
 the same amount of time, with larger chromosome lengths.

Results also show that when increasing the number of islands, the
adaptive overlapping method significantly improves the results with
respect to the disjoint and overlapped methods. However, this is only
apparent with larger chromosome sizes. Therefore, the length of the
chromosome is also a key factor to take into account in adaptive
methods, and not only the number of islands. Studying this factor with
more types of problems, and new configurations of population size and
chromosome lengths will be addressed in the future.% 
% More discussion should be added. Why is adaptive better? Is the high
% performance really needed? - JJ
% STILL more discussion should be needed. For instance, the last
% paragraph of results begs an explanation of the influence of diversity on
% results - JJ

Also, more distributed implementations in several systems (such as a
GPUs or heterogeneous clusters) with different amounts of
islands/processors, will be used to perform a scalability study of the
different methods, being the transmission time between islands a
relevant issue to address. Other MOEA algorithms available in the literature, such as SPEA or MOEA/D will be compared. Finally, new benchmarks and real problems
will also be  used to validate this approach.

\section*{Acknowledgements}
This work has been partially funded by projects TIN2014-56494-C4-3-P y TIN2015-67020-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR),  and  MSTR (PRY142/14, Fundaci\'on P\'ublica Andaluza Centro de Estudios Andaluces en la IX Convocatoria de Proyectos de Investigaci\'on).

>>>>

\begin{abstract}
  % First, state the problem!!!! - JJ Pablo: ok!
%Real-world optimization problems usually require to deal with several objectives.
%Multi-objective optimization methods have gained interest in the last
%years because of the performance they can achieve solving this kind of problems. 
%Many of them usually require a high-dimensional
%decision space, and multi-objective Evolutionary Algorithms have been
%successfully used in the past. Due to the large computational cost
%these algorithms require, different parallelization and distribution
%methods have been proposed, with different parallel and distributed
%computer architectures to cope with them.

One of the most extended
methods to deal with high-dimensional multi-objective optimization problems are the distributed co-evolutionary algorithms. In these methods each
processor deals only with a part of the decision space, that is, the
individuals belonging to the same sub-population (island) explore the
same subset of decision variables. However, only decomposable problems could be strictly implemented in parallel this way. %these methods only works with decomposable problems.

A new approach is presented here to deal with un-decomposable problems, the selective operator application:
instead of sharing parts of individuals, each processor applies the variation 
operators to a specific subset of the whole individual. We have previously proved 
that extending the length of these subsets, and therefore, sharing some search 
space between islands, implies an improvement in the results.

In this paper, a new method to
automatically adapt the size of the overlapping, or shared sections,
of the individuals has been compared to other collaborative
evolutionary techniques. Different number of islands, chromosome sizes
and problems have been used. The analysis of the obtained experimental
results, by using different metrics, shows that our approach can provide statistically significant improvements with
respect to the base algorithm.%, being the method to automatically adapt
%the overlapping size the one to obtain the best performance during the
%same execution time. 
 Results also show that the relation between the number
of islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables. 
\end{abstract}

@@

%2 COMPUTATION -> PARALLEL
These kind of problems implies a high demand of computational time, leading to the apparition of easily parallelizable methods to solve them \citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress}.
%3 ONE WAY IS TO USE MOEAS BECAUSE
One example of easily parallelizable algorithms to solve MOPs are the Multi-Objective Evolutionary Algorithms (MOEAs), a sub-set of Evolutionary algorithms (EAs). EAs \citep{DBLP:series/ncs/EibenS15} are bio-inspired meta-heuristics that can be effectively used to find nearly optimal solutions for optimization problems. Usually an EA starts by generating a set of random solutions, called \emph{population}, following a user-defined description. Then, it evaluates each candidate solution, called \emph{individual}, assigning it a \emph{fitness} value, that describes how good the individual is, with regard to the target problem. New solutions are then generated by the application of \emph{operators} that either mutate a single solution or recombine different existing solutions. After each iteration, called \emph{generation}, the least fit individuals are removed, and the process continues until a user-defined stop condition is met. One of the most well-known and referenced MOEAs \citep{Dorronsoro13superlinear} is the {\em
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, and divided into different ranks (based on the
dominance) to be selected for the next generations.

@@

Different approaches have been used to parallelize EAs, as each individual can be considered as an
independent unit \citep{Alba13parallel}. Classic methods, such as the
global parallel EAs (Master-Slave), or the spatially structured
algorithms (Island model or Cellular EAs) have been applied
successfully in the past \citep{Folino03cellular,Alba02Parallelism}.
However, in the case of MOEAs, these
approaches \citep{Luna15Survey} need to deal with the whole solution set, the PF.
 This implies to use different distribution
and sharing mechanisms, as there exist a tradeoff between the speedup
achievable from parallelization and the need to globally recombine the
results to accurately identify the PF
\citep{Branke04Parallelizingcone}.

MOPs usually require a high number of variables, meaning that the MOEAs need to deal with large individuals and spend significant extra time for crossover, mutation and migration. Different authors have proposed methods to divide the decision space (the chromosome) to improve the performance and solution quality. In this aspect, the co-evolution model is a dimension-distributed model where a high-dimensional problem is divided into lower dimensional ones \citep{Gong15models,Tonda12cooperative}, and evolved separately. One example of application of this technique was described in \citep{Kimovski15Parallel}. The method presented in that works involves different workers that evolve sub-populations created and recombined by a master process, which performs different recombination alternatives from the parts returned by worker processes. A high dimensional problem was used to compare these
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

The hypothesis of this paper is that the use of automatically calculated overlapped sections of the
chromosome for the SLA, can improve the
quality of the solutions in the same computing time with respect to
the baseline, disjoint and partly overlapped methods. 
These overlapped sections are calculated from the number of available islands, in a
multi-objective island-based algorithm, and tested in a real computational environment.

%That was the
                                %hypothesis of the previous
                                %paper. This paper has a new
                                %hypothesis, right? - JJ FERGU: no, this is the hypothesis of this paper (comparing adaptive with partly overlapped), but clarifying it
In order to prove this claim, our new adaptive method has been
compared against different overlapping islands schemes: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
SLA version with disjoint sections and a SLA
fixed-size overlapped version. %What was done in the previous paper?
                               %Are we re-using the results in that
                               %paper? - JJ FERGU: explained before that we are showing new results.
Different benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.  % Shouldn't this start with "We are trying to
                        % prove..." - JJ FERGU: Already writed in previous paragraph

@@

% Is this really cooperative co-evolution? - JJ FERGU: You mean the Dorronsoro one? The title says it
Addressing high-dimensional problems with cooperative co-evolution has
also been studied in several works with approaches closer to the one
presented here. The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. This approach has also been used by Kimovski et al. \citep{Kimovski15Parallel}, but using a
master-slave method that splits the population into several processors. As in previous works, each node
runs a parallel MOEA that only affects some portion of 
the individuals and the master
process receives all the sub-populations to be combined every certain number of generations. Up to 8 processors
were used in the experimental setup, and several
combination alternatives were compared. The main difference of our work with respect to previous works is that our approach does not 
broadcast all solutions to all islands for recombination, but only one solution to a random island, needing less
communication time. Moreover, Dorronsoro or Kimovsky approaches limited the maximum number of islands to 8, while in this paper we have
used up to 128 islands.

@@

In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two different dMOEAs. The first
one divided the chromosome into $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivated us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and this time, the experiments have been run on a parallel
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, and also validating the
previous methods.

@@

%This probably should deserve a "Algorithm" environment too - JJ FERGU: You mean the LaTeX one? I've used it in the pseudocode figures
This basic distributed algorithm spreads the population among $P$
islands; after a fixed number of generations,
one individual in a given island is migrated to another random island,
thus avoiding to synchronize the global Pareto Front (PF) every
certain number of generations as other methods described in Section 2
do. At the end of the run, the Pareto Fronts of all islands are
aggregated into a new one, and the quality measures are evaluated. 
Different  SLA alternatives can be devised to evolve the
subpopulations according to the decision space to be explored by each
islands. Here we will consider three methods in which each island only
performs crossover and mutation in specific sections of the
individuals.

@@

The criterion used for terminating an experiment has been the running
time: 25 seconds for dimension 512 and 100s (four times more) for
2048. We have used time instead of the number of evaluations firstly
because our hypothesis argues that the time saved in crossover and
mutation can be spent on improving the sub-populations and more
operations and migrations can be achieved. Also, we are using
different number of islands (with different sub-population sizes) and that
could lead to different execution times, so it would be difficult to
compare different times and quality of solutions at the same time.
 The whole parameter set is summarized in Table \ref{tab:parameters}.

@@

Different metrics, previously explained, have been used
to calculate the quality of the obtained PFs in each configuration. As
some of the metrics  (such as HV) require a reference point to be
calculated, the point (1,9) has been chosen as reference, as none of
the generated PFs in all runs are dominated by it. Metrics are then
normalized with respect to that point. A Kruskal-Wallis significance
test has been performed to the metrics of all runs of the
configurations, as the Kolmogorov-Smirnov test detected non-normal
distributions. The average results for each configuration are shown in
Table \ref{tab:results512} (for 512 dimensions) and
\ref{tab:results2048} (for 2048). As previously explained, when $P=8$
the results of A are equivalent to the obtained with D (because
$c=0$), and when $P=16$ they are equivalent to O ($c=1$). We refer this equivalence with the word {\em Equiv.} in tables.

@@

%DESCRIPTION OF THE RESULTS
Results show that there is a difference in performance when using 512 or 2048 dimensions. Although in some cases the overlapped methods obtain quality metrics values better or significantly equal than the obtained by the baseline, it is clear that 512 dimensions is not a value high enough to apply these methods. A contradictions arises with respect to the results obtained in our previous work in a shared-memory mono-processor version \citep{Garcia16hpmoon}, where the overlapping method obtained better results than the baseline in both dimension lengths. This can be explained because the added migration latency between nodes of the cluster requires more time than the time saved during the crossover and migration. This was not a problem in the mono-processor version, where the individuals shared the same memory space.

@@

%EXPLANATION OF THE RESULTS
This can be explained comparing the number of non-dominated solutions
in each island and the average number of generations (Table
\ref{tab:sols512} and Table \ref{tab:sols2048}). With both chromosome
lengths, increasing the number of islands implies more generations
with all the methods (logically, as there are less individuals in each
island). But also, the SLA methods
approach to the number of generations with the baseline when
increasing the number of islands. However, with $L=2048$ the migration
time is big enough  to not improve the number of generations
with respect to the baseline, even in the lower number of islands, as
it happens with $L=512$. Therefore, more generations do not
necessarily improve the solution of the global PF, but focusing on
different elements of the chromosome. As indicated previously, every
island is unaware of the other islands PFs, and they are trying to
optimize their solutions independently. With respect to the average
number of solutions in each island, there is a clear difference 
 with the baseline with $L=512$, where this value is in the most
of the cases, less than a half. The number of non-dominated solutions
also implies a better Spread indicator, where the baseline 
 obtains better (or not significantly different) values in the majority of the compared configurations. %TENGO QUE COMENTAR MAS LOS RESULTADOS Y
                              %ENTENDERLO TODO Y JUSTIFICARLO
                              %FINAMENTE
%Almost always no es muy preciso - JJ Pablo: arreglao

@@

Problems that require high-performance % problems are not high-performance. They
                          % _require_ high performance Fergu: done
 and deal with a large number of decision variables can leverage the
 division of the decision space that parallel and distributed
 algorithms imply. This can be done in dMOEAs by selective operator application (SLA), that is, dividing the
 chromosome into different parts, each one modified by a different
 island. This paper compares a baseline distributed NSGA-II with three
 different strategies to separate the chromosome (disjoint or
 overlapping parts), using a high number of islands. Results show that
 these methods can achieve better quality metrics than the baseline in
 the same amount of time, with large chromosome lengths.

Results also show that when increasing the number of islands, the
adaptive overlapping method significantly improves the results with
respect to the disjoint and overlapped methods. However, this is only
apparent with larger chromosome sizes. Therefore, the length of the
chromosome is also a key factor to take into account in adaptive
methods, and not only the number of islands. Studying this factor with
more types of problems, and new configurations of population size and
chromosome lengths could be addressed in the future. For example, comparing 
different ways to calculate $c$ using linear, logarithmic or exponential functions
depending on these values. Moreover, a diversity analysis of the individuals
in each island during the algorithm run could be carried out to understand its influence
on results.
% 
% More discussion should be added. Why is adaptive better? Is the high
% performance really needed? - JJ
% STILL more discussion should be needed. For instance, the last
% paragraph of results begs an explanation of the influence of diversity on
% results - JJ Pablo: extending this section

Also, more distributed implementations in several systems (such as a
GPUs or heterogeneous clusters) with different amounts of
islands/processors, could be used to perform a scalability study of the
different methods, being the transmission time between islands a
relevant issue to be addressed. Other MOEA algorithms available in the literature, such as SPEA or MOEA/D could be compared. Finally, new benchmarks and real problems
could also be  used to validate this approach.

\section*{Acknowledgements}
This work has been partially funded by projects TIN2014-56494-C4-3-P and TIN2015-67020-P (Spanish Ministry of Economy and Competitivity), PROY-PP2015-06 (Plan Propio 2015 UGR),  and  MSTR (PRY142/14, Fundaci\'on P\'ublica Andaluza Centro de Estudios Andaluces en la IX Convocatoria de Proyectos de Investigaci\'on).
